### 2016-02-21 papers/kmeans/response_to_referee.txt
placeholder for referee response

Dear Editor:

Please find below the response to our (very helpful) referee report on
manuscript ApJ/ApJ101852.  We have taken all of the referees
suggestions, and made some other changes from comments we have
received from colleagues.

In what follows, the Referee comments are indented and our responses
are not.  In the revised manuscript, we have made the changes in bold
text so they are apparent to the editor and referee.

     This paper presents new element abundance results using The
     Cannon, a novel, powerful method for spectroscopic abundance
     measurements. The results presented in this paper are new and
     very much relevant to the many large scale observing projects
     that are currently in operation.

     I highlight some comments for the authors to address:

     Abstract: The last sentence is not true, as the authors note
     later in the text the first blind search purely in abundance
     space was carried out by Mitschang et al 2014.

HOGG

     Introduction: The first paragraph is missing a reference or two,
     to support claim that most stars end up in different parts of the
     Galaxy. Or this is something that is assumed and yet to be
     tested?

HOGG

     There are some missing references to the original demonstrations
     of chemical tagging of unbound groups, namely the Wolf 630 group
     and HR 1614 group. Those were key to building up the case for
     this "tagging" research.

HOGG

     The text needs to highlight that the training set is critical -
     Cannon doesn't know physics. It needs to be told what is "true"
     and that the subsequent results will be limited to the believed
     level of truth. While it should be specified that tagging per se
     doesn't require absolute abundances, but the subsequent
     interpretation of the chemical accretion history will require
     "true" abundances - not just precision. Also some clarification
     about the Cannon's performance at the boundaries of the training
     set parameters would have been good to see.

HOGG

     Section 2:

     A reference for the famously demonstrated correlations in
     globular clusters is needed.

HOGG

     Section 3:

     The text doesn't clarify why only k-means was chosen to do the
     'tagging' given the issues discussed, was it simply easy and
     fast? Fine as the focus of the paper is the precision of the
     abundance results with Cannon, not the fidelity of group finding.

HOGG

     I find the group in Fig 6 dubious. Its at the very edge of the
     metallicity scale, which raises again how the Cannon performs at
     the boundaries of the training parameters. Could this not simply
     be metal-poor Halo stars, tagged as metal-poor Halo stars,
     nothing more?

HOGG

-----
Other changes:

In addition to all the above, we made the following additional
changes:

HOGG:  We added authors!

HOGG:  We added citations!

HOGG:  Adjusted what we say about Ting et al?

HOGG:  Go through github issues and emails.