### 2016-03-22 papers/nucleosynthesis/nuke.tex
placeholder

\documentclass[12pt, preprint]{aastex}

\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\apogee}{\project{\acronym{APOGEE}}}

\newcommand{\solar}{\odot}

\begin{document}

\title{The minimal nucleosynthesis project with \apogee}
\author{DWH, HWR, JR, MKN, ARC, others}

\begin{abstract}
We describe the minimal thing you could do with the \apogee\ data to
improve our understanding of nucleosynthesis.
\end{abstract}

\keywords{
  ---
  World: Hello
  ---
}

\section{Introduction}

Nucleosynthesis is important.  Models are wrong.  Data are now abundant.

\section{Model}

\begin{eqnarray}
  [X/Y] &\equiv& \log_{10}\left(\frac{X}{Y}\right) -
                 \log_{10}\left(\frac{X}{Y}\right)_\solar
\end{eqnarray}

\end{document}