### 2016-03-23 papers/nucleosynthesis/nuke.tex
model in equations

\documentclass[12pt, preprint]{aastex}

% text commands
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\apogee}{\project{\acronym{APOGEE}}}

% math commands
\newcommand{\solar}{\odot}
\newcommand{\abundance}[1]{\lbrack{#1}\rbrack}
\newcommand{\given}{\,|\,}
\newcommand{\normal}{\mathcal{N}}

\begin{document}

\title{The minimal nucleosynthesis project with \apogee}
\author{DWH, HWR, JR, MKN, ARC, others}

\begin{abstract}
We describe the minimal thing you could do with the \apogee\ data to
improve our understanding of nucleosynthesis.
\end{abstract}

\keywords{
  ---
  World: Hello
  ---
}

\section{Introduction}

Nucleosynthesis is important.  Models are wrong.  Data are now abundant.

\section{Model}

\begin{eqnarray}
  \abundance{X/Y}_n &\equiv& \log_{10}\left(\left.\frac{X}{Y}\right|_n\right) -
                             \log_{10}\left(\left.\frac{X}{Y}\right|_\solar\right)
\\
  \abundance{X/Y}_n &=& \log_{10}(y_{nd}) + e_{nd}
\\
  p(e_{nd}) &=& \normal(e_{nd}\given 0,\sigma^2_{nd})
\\
  y_{nd} &=& \sum_{k=1}^K a_{nk}\,g_{kd}
\\
  g_{kd} &=& G_{kd} + \epsilon_{kd}
\\
  p(\epsilon_{kd}) &=& \normal(\epsilon_{kd}\given 0,\Sigma^2_{kd})
\end{eqnarray}

\end{document}