### 2016-04-15 papers/kmeans/response_to_referee.txt

Dear Editor:

Please find below the response to our (very helpful) referee report on
manuscript ApJ/ApJ101852.  We have taken all of the referees
suggestions, and made some other changes from comments we have
received from colleagues.

In what follows, the Referee comments are indented and our responses
are not.  In the revised manuscript, we have made the changes in bold
text so they are apparent to the editor and referee.

    This paper presents new element abundance results using The
    Cannon, a novel, powerful method for spectroscopic abundance
    measurements. The results presented in this paper are new and
    very much relevant to the many large scale observing projects
    that are currently in operation.

    I highlight some comments for the authors to address:

    Abstract: The last sentence is not true, as the authors note
    later in the text the first blind search purely in abundance
    space was carried out by Mitschang et al 2014.

        We have referenced Mitschang et al. (2014) for their work in
        chemical tagging, but we believe the last sentence of the
        abstract is correct. To clarify: Mitschang et al. (2014) did
        perform a blind search in chemical abundances to identify
        substructure, but one of the key conclusions of their work
        was that ".. these groups may not represent distinct 
        formation sites, e.g. as in dissolved open clusters, as 
        previously thought". While the Mitschang et al. (2014) work
        showed their chemically-identified structures had similar 
        (U, V, W) space velocities (e.g., their Figure 2), those 
        velocities are not dissimilar from the background velocity
        distribution. In contrast, our work is the first blind
        search in chemical abundances that has identified structures
        that are *also* structures in phase-space: specifically
        on-sky position and radial velocities.

        Nevertheless, to avoid unwarranted claims, we have qualified
        this statement from:

        "This is the first-ever project to identify phase-space 
        structures by blind search purely in abundance space;
        the prospects for future data sets are very good."

        to:

        "This is the first-ever project to identify phase-space 
        structures at survey-scale by blind search purely in abundance space;
        the prospects for future data sets are very good."

        since the Mitschang et al. (2014) work had a sample size of
        714 stars.

    Introduction: The first paragraph is missing a reference or two,
    to support claim that most stars end up in different parts of the
    Galaxy. Or this is something that is assumed and yet to be
    tested?

        The text claims that stars end up in different parts of
        phase-space in the Galaxy: that is to say that any common
        kinematic (or dynamic: actions, angles) signature between
        stars born in the same molecular gas cloud at the same time
        will eventually disperse, presumably throughout the disk.
        This line of reasoning follows from expectations that most
        stars form in clusters, but the disk is full of stars that
        do not appear to be kinematically bound to a cluster. For
        these reasons we would claim that the opening paragraph is
        more of an introduction to the expectations from star
        formation, or rather introducing the line of reasoning for
        why elemental abundances can be instructive for reconstructing
        the evolution of the Galaxy.

    There are some missing references to the original demonstrations
    of chemical tagging of unbound groups, namely the Wolf 630 group
    and HR 1614 group. Those were key to building up the case for
    this "tagging" research.

        We have added the following references: De Silva et al. 
        (2007), Eggen (1970), and Bubar & King (2010).

    The text needs to highlight that the training set is critical -
    Cannon doesn't know physics. It needs to be told what is "true"
    and that the subsequent results will be limited to the believed
    level of truth. While it should be specified that tagging per se
    doesn't require absolute abundances, but the subsequent
    interpretation of the chemical accretion history will require
    "true" abundances - not just precision. Also some clarification
    about the Cannon's performance at the boundaries of the training
    set parameters would have been good to see.

HOGG

    Section 2:

    A reference for the famously demonstrated correlations in
    globular clusters is needed.

        We have added references to: Norris & Da Costa (1995), which
        (to our knowledge) was the first to unambiguously show anti-
        correlations in Na, Al, and O, and to Carretta et al. (2009),
        which provides a more recent and comprehensive view of our
        current knowledge of elemental abundance correlations in
        gloublar clusters.

    Section 3:

    The text doesn't clarify why only k-means was chosen to do the
    'tagging' given the issues discussed, was it simply easy and
    fast? Fine as the focus of the paper is the precision of the
    abundance results with Cannon, not the fidelity of group finding.

        We have added text to clarify why k-means was chosen.
        Specifically we have updated the paragraph:

        "Nothing about this algorithm or our choices are particularly
        tuned or optimized; this is in no sense the
        algorithm or the method that reveals the best structures.
        Indeed, it is a feature of this work that even the simplest, most
        generic clustering algorithm returns interesting
        structures (as shown below). The prospects for Milky Way science only 
        improve as the clustering algorithm improves."

        To now read as:

        "Nothing about this algorithm or our choices are particularly
        tuned or optimized; this is in no sense the
        algorithm or the method that reveals the best structures.
        We chose k-means as a simple and tractable approach to identifying
        clusters in high dimensional label space. In practice more complex
        clustering algorithms may perform better than k-means, or even
        tuning k-means by heuristically setting $K$ is likely to improve
        upon the results here. This \documentname\ serves as a proof of concept.
        Indeed, it is a feature of this work that even the simplest, most
        generic clustering algorithm returns interesting
        structures (as shown below). The prospects for Milky Way science only 
        improve as the clustering algorithm improves."

    I find the group in Fig 6 dubious. Its at the very edge of the
    metallicity scale, which raises again how the Cannon performs at
    the boundaries of the training parameters. Could this not simply
    be metal-poor Halo stars, tagged as metal-poor Halo stars,
    nothing more?

HOGG

-----
Other changes:

In addition to all the above, we made the following additional
changes:

Because of a large number of valuable contributions both prior to and
subsequent to our submission of this paper, we expanded the author
list.  There are now several additional new authors, most of whom
contributed by building the APOGEE project (and hence the input data).

We added a few citations: Eisenstein 2011 (SDSS-III), Holtzman 2015
(APOGEE DR12) and ...HOGG

HOGG:  Adjusted what we say about Ting et al?

HOGG:  Go through github issues and emails.

We also added a few names to the acknowledgements for providing minor
comments.