### 2016-08-06 papers/kmeans/response_to_referee.txt
authors, affils, response text

Dear Editor:

Please find below the response to our (very helpful) referee report on
manuscript ApJ/ApJ101852.  We have taken all of the referee's
suggestions, and made some other changes from comments we have
received from colleagues.

In what follows, the Referee comments are indented and our responses
are not.  In the revised manuscript, we have made the changes in bold
text so they are apparent to the editor and referee.

    This paper presents new element abundance results using The
    Cannon, a novel, powerful method for spectroscopic abundance
    measurements. The results presented in this paper are new and
    very much relevant to the many large scale observing projects
    that are currently in operation.

    I highlight some comments for the authors to address:

    Abstract: The last sentence is not true, as the authors note
    later in the text the first blind search purely in abundance
    space was carried out by Mitschang et al 2014.

We have referenced Mitschang et al. (2014) for their work in chemical
tagging, but we believe the last sentence of the abstract is
correct. To clarify: Mitschang et al. (2014) did perform a blind
search in chemical abundances to identify substructure, but one of the
key conclusions of their work was that ".. these groups may not
represent distinct formation sites, e.g. as in dissolved open
clusters, as previously thought". While the Mitschang et al. (2014)
work showed their chemically-identified structures had similar (U, V,
W) space velocities (e.g., their Figure 2), those velocities are not
dissimilar from the background velocity distribution. In contrast, our
work is the first blind search in chemical abundances that has
identified structures that are *also* structures in phase-space:
specifically on-sky position and radial velocities.

Nevertheless, to avoid unwarranted claims, we have qualified this
statement in the abstract.

    Introduction: The first paragraph is missing a reference or two,
    to support claim that most stars end up in different parts of the
    Galaxy. Or this is something that is assumed and yet to be
    tested?

The text claims that stars end up in different parts of phase-space in
the Galaxy: that is to say that any common kinematic (or dynamic:
actions, angles) signature between stars born in the same molecular
gas cloud at the same time will eventually disperse, presumably
throughout the disk.  This line of reasoning follows from expectations
that most stars form in clusters, but the disk is full of stars that
do not appear to be kinematically bound to a cluster. For these
reasons we would claim that the opening paragraph is more of an
introduction to the expectations from star formation, or rather
introducing the line of reasoning for why elemental abundances can be
instructive for reconstructing the evolution of the Galaxy.

    There are some missing references to the original demonstrations
    of chemical tagging of unbound groups, namely the Wolf 630 group
    and HR 1614 group. Those were key to building up the case for
    this "tagging" research.

We have added the following references: De Silva et al (2007), Eggen
(1970), and Bubar & King (2010).

    The text needs to highlight that the training set is critical -
    Cannon doesn't know physics. It needs to be told what is "true"
    and that the subsequent results will be limited to the believed
    level of truth. While it should be specified that tagging per se
    doesn't require absolute abundances, but the subsequent
    interpretation of the chemical accretion history will require
    "true" abundances - not just precision. Also some clarification
    about the Cannon's performance at the boundaries of the training
    set parameters would have been good to see.

This statement and point is true and wise. We have modified the text
in [HOGG]

    Section 2:

    A reference for the famously demonstrated correlations in
    globular clusters is needed.

We have added references to: Norris & Da Costa (1995), which (to our
knowledge) was the first to unambiguously show anti- correlations in
Na, Al, and O, and to Carretta et al. (2009), which provides a more
recent and comprehensive view of our current knowledge of elemental
abundance correlations in gloublar clusters.

    Section 3:

    The text doesn't clarify why only k-means was chosen to do the
    'tagging' given the issues discussed, was it simply easy and
    fast? Fine as the focus of the paper is the precision of the
    abundance results with Cannon, not the fidelity of group finding.

We have added text to clarify why k-means was chosen. It is
fundamentally because k-means is the clustering algorithm of "first
resort", which contains very few assumptions and only one (integer)
adjustable parameter (not counting the metric on the vector space).

    I find the group in Fig 6 dubious. Its at the very edge of the
    metallicity scale, which raises again how the Cannon performs at
    the boundaries of the training parameters. Could this not simply
    be metal-poor Halo stars, tagged as metal-poor Halo stars,
    nothing more?

HOGG

-----
Other changes:

In addition to all the above, we made the following additional
changes:

Because of discussions and checking of results that happened
subsequent to our submission of this paper with the APOGEE team, some
of which made use of proprietary data and information about the
survey, we have expanded the author list in accordance with the
APOGEE/SDSS-III authorship policies.  The consequence is that there
are now several additional new authors, most of whom contributed to
this project most directly by building the APOGEE project (and hence
the input data and tests thereof).

We added the two crucial APOGEE references: Eisenstein 2011 (SDSS-III)
Holtzman 2015 (APOGEE DR12).

HOGG:  Adjusted what we say about Ting et al?

HOGG:  Commented on, eg, the Na abundances?

HOGG:  Go through github issues and emails?

We also added a few names to the acknowledgements for providing minor
comments.